Sketch Recognition using Particle Swarm Optimizations

### 2009-06-10 paper2/Paper2.tex
figure comment removal
after chanign and addin omega image

\section{Discrete Particle Swarm Optimization}
\label{sec:ParticleSwarmAlgorithm}
%\section{Particle Swarm Algorithm}
%\label{PSO}
%What is particle swarm algorithm and how it was used in related researches. 
%The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution in the solution space with a particle\cite{PSOFirst}. Each particle moves the n with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.
The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}. Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.

\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}
%where $p_{ij}$ represent the $jth$ particle in the $ith$ agent and $v_{ij}$ is the velocity of the $jth$ particle in the $ith$ agent.
where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
\label{eq:Swarm}
\end{equation}
 where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.

@@

After each iteration of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled, the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, two temperary particles are evaluated by removing either point from the current particle. After that, one of these points is removed based on the error caused by its removal.
	\item Each two successive segments are tested to determine if they can be merged into a single segment. If merging the two segments results in segmentation that has an error value less than or equal to the error before merging, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
	% the same or smaller the error
\end{enumerate}
%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The following list describes each feature in details.
%\begin{description}
\subsubsection{Structural and geometrical Features (FS1)}
  They are features that define the structure of the geometrical symbol such as;   
%\begin{description}
%	 \item[No. Of segments] Number of segments in the symbol. 
%		\item[No. Of primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [No. of curves] Number of curves or ellipses in the symbol. 
%		\item [No. of lines] Number of lines in the symbol. 
%		\item [No. of perpendicular lines] Number of perpendicular lines. 
%		\item [No. of parallel lines] Number of parallel lines. 
%		\item [No. of intersections] Number of intersection between lines and curves. 
%		\item [Size Ratio] Ratio between width to height of the symbol.
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
 \begin{itemize}
	 \item \emph{Segments:} The number of segments in the symbol.
	 \item \emph{Strokes:} The number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} The number of primitives in the symbol. This feature helps when identifying  symbols with mixed geometric primitives such as cylinders and callouts.  
		\item \emph{Curves:} The number of curves or ellipses in the symbol. 
		\item \emph{Lines:} The number of lines in the symbol. 
		\item \emph{Perpendicular}\emph{lines:} The number of perpendicular lines. 
		\item \emph{Parallel}\emph{lines:} The number of parallel lines. 
		\item \emph{Intersections:} The number of intersection between lines and curves. 
		\item \emph{T intersections:} The number of T intersections. 
		\item \emph{L intersections:} The number of L intersections. 
		\item \emph{X intersections:} The number of X intersections.	
		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
\subsubsection{Rubine Features (FS2)}
  These features were introduced by Rubine \cite{gestureexample12} for single stroke gestures. Such as length of stroke, cosine of starting angle, smoothness of symbol and some features extracted from the bounding box of the symbol.

\subsubsection{Statistical Features (FS3)} 
These features use moments and statistical properties such as;
\begin{itemize}
	\item \emph{Zernike moments:} The zernike moments are orthogonal moments which define the distribution of point in the input space. They are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}. %The zernike moment were used in \cite{HeloiseBeautification} t. 
\end{itemize}
\subsubsection{Composite Features (FS4)}
 These features are composites of different geometrical and statistical properties of the symbol. 
	\begin{itemize}
\item \emph{Size Ratio:} The ratio between width and height of the symbol.
	\item \emph{Ink density:} The density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item \emph{Convex Hull Area:} The area of convex hull with respect to area of bounding box of symbol.
	\item \emph{Convex Hull Perimeter:} The perimeter of convex hull with respect to total length of symbol.
		\item \emph{Mean Centroidal radius:} The Mean of the centroidal radius is the distance from each point in the symbol to the center of gravity.
	\item \emph{Mean Time difference:} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]
  \end{itemize}  
%\end{description}
%The features used consist of Rubine feature set,  of order 21 , ink density as well as some structural and spatial information like number of perpendiculars lines ,number of parallel lines and number of different types of primitives in each symbol. 
After computing the features, the symbol is introduced to the classifier as a feature vector. The system uses Support Vector Machine (SVM) classifier with Linear kernel \cite{libsvm}. An OVO classifier structure (object versus object) is used to handle multi-class problem.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.

@@

%The dataset collected by Hse and Newton\cite{HeloiseBeautification} is used to test the presented system. The data are drawn by 16 users each of them was asked to draw 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. The dataset was divided into training set and test set. Four different splits for the training and test data are generated from the dataset. The results displayed are the average recognition accuracy of the four splits. The accuracy is computed as the number of correctly recognized samples divided by the total number of samples in the test. %In the following section, the recognition accuracy means the average recognition accuracy in the four data splits.

A set of tests were performed on the system to enhance non-ellipse fitting algorithms. The effect of the number of particles, the error thresholds and various other parameters were investigated. The results of these tests are shown in Figures \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}. Figure \ref{fig:swarmtesting} shows that as the number of swarm iterations increase the error value decrease (error of the segmentation algorithm). The final number of particles and maximum swarm iterations used in the system was based on a tradeoff between error achieved versus the computation time. The effect of other algorithm parameters such as $c_1$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure \ref{fig:swarmtesting}, \ref{fig:swarmtesting2}).%, $c_2$

 \begin{figure}
	\centering		
	 	\subfigure[Segmentation Error Vs. Swarm Iterations]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}
	 %\includegraphics[scale=0.3]{images/swarmAgentTest.jpg}} 	\caption{}

	%\caption{Experiments results :} a)    b) c) 
	%\end{figure}

%\begin{figure}
	%\centering

		\subfigure[Segmentation Error Vs. c1]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.4]{images/swarmParamterC.jpg}}

	 	\caption{Segmentation Error Vs. PSO parameters.}a) The error decrease as the number of swarm iteration increase until it reach a saturation.   b) The behavior of segmentation error versus c1.

	%\Segmentation Error Vs. PSO parameters. caption{Experiments results :} a)    b) c)

\end{figure} 
\subsection{Algorithms comparisons}
\label{sec:AlgExp}
We performed several experiments to evaluate the presented recognition system. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms. Figure \ref{fig:test1} and \ref{fig:testLD} show the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. Combining the ellipse detection module improves the performance in certain datasets. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1 in all datasets.  The results also shows that combining (\textsl{AlgS1} and \textsl{AlgS2} using minimum error) outperforms AlgS2 in Hs-Ds only. This behavior is explained by noticing that features are computed based on the chosen segmentation. This means that different segmentation for a symbol generates a different set of features used for recognition. This behavior affects Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure nature of the symbols in both datasets than of Hs-Ds.  %   can be explained by the simplicity of the symbols in Hs-DS % and LD-Ds unlike EL-DS where AlgS2 is the best algorithm. This behavior is explained by the structure of the symbols in each dataset. El-DS consist of highly topological and structure symbols that are usually drawn using multiple small strokes. This leads to    %   out perform any single algorithm in HS-DS and LD-DS.  %show that combining algorithms achieve higher accuracy than any single algorithm. 
%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.

 \begin{figure}[h]
	\centering		
	\subfigure[Algorithm comparison on HS-DS]{\label{fig:test1} \includegraphics[scale=0.4]{images/testAlg.jpg}}
	 %	\caption{} The recognition rate of different algorithms.

	%\caption{Experiments results :} a)    b) c) 
%\end{figure}

%\begin{figure}[h]
%	\centering		
		\subfigure[Algorithm comparison on LD-DS]{ 	 	\label{fig:testLD} \includegraphics[scale=0.4]{images/LDAlg.jpg}}
	% 	\caption{} The recognition rate of different algorithms.

	%\caption{Experiments results :} a)    b) c) 
%\end{figure}

%\begin{figure}[h]
%	\centering		
		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgEL.jpg}}
	 	\caption{ The recognition rate of different algorithms. }

	%\caption{Experiments results :} a)    b) c) 
\end{figure}

@@

\begin{figure*}[h]

@@

	%\caption{} The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
%\end{figure*}  
%\begin{figure*}
%	\centering

@@

 \begin{figure*}[h]
	\centering
	\subfigure[Feature Comparison on Hs-DS]{	\label{fig:testFeaturesAll} \includegraphics[scale=0.5]{images/featuresdifference.jpg}}
    %The graph shows the recognition rate of each symbol using different algorithms.

%\end{figure*}

 %\begin{figure*}
	%\centering

	\subfigure[Feature Comparison on EL-DS]{	\label{fig:ELtestFeaturesAll}	\includegraphics[scale=0.34]{images/FeatEl.jpg}}
	\caption{Different Features} The effect of different feature set on accuracy.  FS4 and FS3 give good results in a) Hs-DS and b) EL-Ds. The results are improved by combining FS1 with FS2 and FS1 with FS4.    %The graph shows the recognition rate of each symbol using different algorithms. 
\end{figure*}

@@

 The main features affected by the segmentation stage are the geometric features computed in \textbf{FS1}. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly}. The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1} features. 
\begin{figure}[h] 
	\centering
		\includegraphics[scale=0.3]{images/OnlyF1EL.jpg}
	\caption{The effect of segmentation on accuracy} The experiment is implemented on Hs-Ds using only FS1.  %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}

@@

This paper presented a new approach to sketch recognition using DPSO. The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}. It is noted that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results. The final recognition is based on both global shape and hierarchical stroke properties. Adding stroke and segment geometrical features improved the final recognition. Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}. The experiments evaluated two\textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds). Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the datasets used. The results proved that the system is easily expandable to more and more symbols as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.  The system achieved an overall improvement of more than 2\% over Alg3 in all datasets.  %The system was tested on 13 different symbols and achieved an overall accuracy of 89.5\%. The system does not depend on low level recognizer but rather on a set of high level features. This makes the system easily expandable to more and more symbols.

 A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. Currently, the system processes one symbol at a time, an enhancement would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols. Another area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications. Features that represents the appearance of the symbols are likely to improve the recognition that have similar geometrical structure (i.e. Nor and OR gate in logic symbols).  
 \bibliographystyle{elsarticle-harv} 
%\bibliographystyle{IEEEbib}
\bibliography{../../neededfiles/Bibliographies/Mybibliography}
\end{document}

>>>>

\section{Discrete Particle Swarm Optimization}
\label{sec:ParticleSwarmAlgorithm}
 The main idea of \textit{Particle Swarm Algorithm (PSO)} is to represent each solution with a $N$ dimension particle from the solution space \cite{PSOFirst}. Each particle moves with a direction and velocity $v_{ij}$ based on equations \ref{eq:Swarm1} \& \ref{eq:Swarm}.

\begin{equation}
%\[
p_{ij}=p_{ij}+v_{ij},
%\
\label{eq:Swarm1}
\end{equation}

where $p_{ij}$ represent the $j$th dimension in the $i$th particle and $v_{ij}$ is the velocity of the $j$th dimension in the $i$th particle.
 %Equation [\ref{eq:Swarm}] shows how velocity and direction of each particle are computed
 \begin{equation}
v_{ij}  = v_{ij} \omega + c_1 r_1 (lbest_{ij}  - p_{ij} ) + c_2 r_2 (gbest_{ij}  - p_{ij} ),
\label{eq:Swarm}
\end{equation}
 where $\omega$ is the inertia weight parameter which controls the tradeoff between exploration and exploitation,  $lbest_{ij}$ is the local best particle, $gbest_{ij}$ is the global best particle, $r_1$ \& $r_2$ are random variables and $c_1$ \& $c_2$ are the swarm acceleration parameters.

@@

After each iteration of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedures: 
\begin{enumerate}
	\item For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{Prepross}). If it was not labeled, the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. 
	\item The particles are tested to make sure that the distance between every two successive dominate point is larger than $min_D$, where $min_D$ is 5\% of the total length of the stroke.  Otherwise, two temperary particles are evaluated by removing either point from the current particle. After that, one of these points is removed based on the error caused by its removal.
	\item Each two successive segments are tested to determine if they can be merged into a single segment. If merging the two segments results in segmentation that has an error value less than or equal to the error before merging, the two segments are merged. This step ensures that the final segmentation of the system has minimum number of segments. 
	% the same or smaller the error
\end{enumerate}
%1) 2)  % If two points are nearer than $min_D$ then one of the points is removed. 
\subsection{Recognition}
\label{sec:Recognition}
After the user draws all strokes of the symbol, the set of un-recognized strokes is grouped together along with their segmentation as input to the feature extraction process. A composite set of features is used to generate a single feature vector. The following list describes each feature in details.
%\begin{description}
\subsubsection{Structural and geometrical Features (FS1)}
  They are features that define the structure of the geometrical symbol such as;   
 \begin{itemize}
	 \item \emph{Segments:} The number of segments in the symbol.
	 \item \emph{Strokes:} The number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} The number of primitives in the symbol. This feature helps when identifying  symbols with mixed geometric primitives such as cylinders and callouts.  
		\item \emph{Curves:} The number of curves or ellipses in the symbol. 
		\item \emph{Lines:} The number of lines in the symbol. 
		\item \emph{Perpendicular}\emph{lines:} The number of perpendicular lines. 
		\item \emph{Parallel}\emph{lines:} The number of parallel lines. 
		\item \emph{Intersections:} The number of intersection between lines and curves. 
		\item \emph{T intersections:} The number of T intersections. 
		\item \emph{L intersections:} The number of L intersections. 
		\item \emph{X intersections:} The number of X intersections.

\end{itemize}
\subsubsection{Rubine Features (FS2)}
  These features were introduced by Rubine \cite{gestureexample12} for single stroke gestures. Such as length of stroke, cosine of starting angle, smoothness of symbol and some features extracted from the bounding box of the symbol.

\subsubsection{Statistical Features (FS3)} 
These features use moments and statistical properties such as;
\begin{itemize}
	\item \emph{Zernike moments:} The zernike moments are orthogonal moments which define the distribution of point in the input space. They are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}. %The zernike moment were used in \cite{HeloiseBeautification} t. 
\end{itemize}
\subsubsection{Composite Features (FS4)}
 These features are composites of different geometrical and statistical properties of the symbol. 
	\begin{itemize}
\item \emph{Size Ratio:} The ratio between width and height of the symbol.
	\item \emph{Ink density:} The density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item \emph{Convex Hull Area:} The area of convex hull with respect to area of bounding box of symbol.
	\item \emph{Convex Hull Perimeter:} The perimeter of convex hull with respect to total length of symbol.
		\item \emph{Mean Centroidal radius:} The Mean of the centroidal radius is the distance from each point in the symbol to the center of gravity.
	\item \emph{Mean Time difference:} The mean of the time difference between each two successive points in the symbol. 
  \end{itemize}

 After computing the features, the symbol is introduced to the classifier as a feature vector. The system uses Support Vector Machine (SVM) classifier with Linear kernel \cite{libsvm}. An OVO classifier structure (object versus object) is used to handle multi-class problem.% A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$.

@@

A set of tests were performed on the system to enhance non-ellipse fitting algorithms. The effect of the number of particles, the error thresholds and various other parameters were investigated. The results of these tests are shown in Figures \ref{fig:swarmtesting} and   \ref{fig:swarmtesting2}. Figure \ref{fig:swarmtesting} shows that as the number of swarm iterations increase the error value decrease (error of the segmentation algorithm). The final number of particles and maximum swarm iterations used in the system was based on a tradeoff between error achieved versus the computation time. The effect of other algorithm parameters such as $c_1$, $\omega$ (mentioned in Section\ref{sec:ParticleSwarmAlgorithm}) is tested using similar tests (Figure  \ref{fig:swarmtesting2},\ref{fig:swarmwtest}).%, $c_2$

 \begin{figure}
	\centering		
	 	\subfigure[Segmentation Error Vs. Swarm Iterations]{	\label{fig:swarmtesting} \includegraphics[scale=0.3]{images/swarmIterationVsParticles.jpg}}

		\subfigure[Segmentation Error Vs. $c_1$]{	\label{fig:swarmtesting2}  \includegraphics[scale=0.3]{images/C1Swarm.jpg}} %swarmParamterC.jpg

				\subfigure[Segmentation Error Vs. $\omega$]{\label{fig:swarmwtest}  \includegraphics[scale=0.3]{images/wSwarm.jpg}} %swarmParamterC.jpg

	 	\caption{Segmentation Error Vs. PSO parameters.}a) The error decrease as the number of swarm iteration increase until it reach a saturation.   b) The behavior of segmentation error versus $c_1$. c) The behavior of segmentation error versus  $\omega$

\end{figure} 
\subsection{Algorithms comparisons}
\label{sec:AlgExp}
We performed several experiments to evaluate the presented recognition system. Firstly, we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to compare it with our swarm algorithms. Figure \ref{fig:test1} and \ref{fig:testLD} show the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. Combining the ellipse detection module improves the performance in certain datasets. The result shows that (\textsl{AlgS2}) gives better performance than Alg3 and AlgS1 in all datasets.  The results also shows that combining (\textsl{AlgS1} and \textsl{AlgS2} using minimum error) outperforms AlgS2 in Hs-Ds only. This behavior is explained by noticing that features are computed based on the chosen segmentation. This means that different segmentation for a symbol generates a different set of features used for recognition. This behavior affects Ls-Ds and EL-Ds more than Hs-Ds due to the higher topological and structure nature of the symbols in both datasets than of Hs-Ds.

 \begin{figure} 
	\centering		
	\subfigure[Algorithm comparison on HS-DS]{\label{fig:test1} \includegraphics[scale=0.4]{images/testAlg.jpg}}
 		\subfigure[Algorithm comparison on LD-DS]{ 	 	\label{fig:testLD} \includegraphics[scale=0.4]{images/LDAlg.jpg}}
 		\subfigure[Algorithm comparison on EL-DS]{	 	\label{fig:testEL}  \includegraphics[scale=0.4]{images/AlgEL.jpg}}
	 	\caption{ The recognition rate of different algorithms. }
 \end{figure}

@@

\begin{figure*} 

@@

@@

 \begin{figure*}
	\centering
	\subfigure[Feature Comparison on Hs-DS]{	\label{fig:testFeaturesAll} \includegraphics[scale=0.5]{images/featuresdifference.jpg}}
	\subfigure[Feature Comparison on EL-DS]{	\label{fig:ELtestFeaturesAll}	\includegraphics[scale=0.34]{images/FeatEl.jpg}}
	\caption{Different Features} The effect of different feature set on accuracy.  FS4 and FS3 give good results in a) Hs-DS and b) EL-Ds. The results are improved by combining FS1 with FS2 and FS1 with FS4.    %The graph shows the recognition rate of each symbol using different algorithms. 
\end{figure*}

@@

 The main features affected by the segmentation stage are the geometric features computed in \textbf{FS1}. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly}. The figure shows that \textsl{AlgS2} achieve better results.  
\begin{figure*} 
	\centering
		\includegraphics[scale=0.45]{images/OnlyF1EL.jpg}
	\caption{The effect of segmentation on accuracy} The experiment is implemented on Hs-Ds using only FS1.  %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure*}

@@

This paper presented a new approach to sketch recognition using DPSO. The system uses multiple preliminary information, i.e., speed and curvature information, as input to the \textit{DPSO}. Results show that the \textit{DPSO} in general generates an optimized stroke segmentation which improves the final recognition rate.  The final recognition is based on both global shape and hierarchical stroke properties. Adding stroke and segment geometrical features improved the final recognition. Global shape and moment properties were used to prevent the need to use graph matching which is computationally expensive \cite{SketchRead2007}. The experiments evaluated two \textit{DPSO} different segmentation algorithms (AlgS1 and AlgS2) on three different datasets (Hs-Ds, EL-Ds and LD-Ds). Results show that AlgS2 achieves better segmentation and final accuracy than Alg3 \cite{earlyprocess} on the datasets used. The results proved that the system is easily expandable to more and more symbols as it does not depend on low level primitive recognizers but on high level symbol and geometrical features.  The system achieved an average overall improvement of more than 2\% over Alg3 in all datasets.

  The tradeoff between accuracy achieved and time complexity can be further investigated to achieve better results. Currently, the system processes one symbol at a time, an enhancement would be to allow users to draw more than one symbol at a time and incorporate a method for separating symbols. Thus, possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. Another area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications. Features that represent the appearance of the symbols are likely to improve the recognition of symbols that have similar geometrical structure (i.e. Nor and OR gate in logic symbols).  
 \bibliographystyle{elsarticle-harv} 
%\bibliographystyle{IEEEbib}
\bibliography{../../neededfiles/Bibliographies/Mybibliography}
\end{document}
